---
title: "Analysis of tree frog limb measurements"
author: "Sara Gama"
date: "2025-01-07"
output: html_document
---

My GitHub repository is accessible here (<https://github.com/SaraGama4/BIOL432_Assignment1_SG>).

## 1: Loading files

Load required libraries
```{r}
library(dplyr)
library(ggplot2)
library(tidyr)
```

I loaded the revised 'measurements.csv' file which now includes the 'Volume' column. I also loaded in my generated data and estimated volume scripts.

In the generated data script I chose five Canadian tree frog species and put them in a vector which I then randomly sampled from for a total of n= 100. I did the same process for generating three recorded names from observations. As for generating hypothetical values for limb width and lengths, I set plausible parameters for the normal distribution including a mean and standard deviation for both variables. This was followed by generating 100 random values from said normal distribution. Finally I combined all of the vectors to be separate columns in a csv file!

In the estimated volume script I read the csv file into a data frame and estimated limb volume using the formula for the volume of a cylinder (for biological relevance). I then mutated the values for each calculation into rows of a new column labelled "volume".

```{r}
source("C:/Users/sgama/Downloads/432/BIOL432_Assignment1_SG/dataGenerato.R")

source("C:/Users/sgama/Downloads/432/BIOL432_Assignment1_SG/volumeEstimato.R")

read.csv("C:/Users/sgama/Downloads/432/BIOL432_Assignment1_SG/measurements.csv")

```

## 2: Sorting the data

I sorted the data first by species, then by observer, and by the limb volume using the arrange function

```{r}
sorted_df <- measurements_df %>%
  arrange(random_species, random_observer, Volume)

head(sorted_df)
```

## 3: Calculating averages for a variable of interest

I calculated the average limb volume for each species in a table

```{r}
avg_vol_species <- measurements_df %>%
  group_by(random_species) %>%
  summarize(avg_vol = mean(Volume, na.rm = TRUE))

avg_vol_species

```

## 4: Counting the number of observations occurring in two variables

I made a table for the number of observations (rows) for each combination of species and observer.

```{r}
obs_species_observer <- measurements_df %>%
  group_by(random_species, random_observer) %>%
  summarise(num_obs = n())

obs_species_observer
```

## 5: Visualizing tree frog limb volume distributions

I created a box plot comparing limb volumes for each species

```{r, fig.dim = c(8, 5)}
ggplot(measurements_df, aes(x = random_species, y = Volume)) +
  geom_boxplot(aes(fill = random_species)) +
  labs(title = "Limb Volumes by Species",
       x = "Tree Frog Species",
       y = "Limb Volume (cm³)") +
  theme_light()
```

## 6: Visualizing distributions of limb volumes in a histogram

I created a multi-panel plot of frequency histograms by species

```{r}
ggplot(measurements_df, aes(x = Volume)) +
  geom_histogram(aes(fill = random_species),binwidth = 3) +
  facet_wrap(~random_species) +
  labs(title = "Distribution of Limb Volumes by Species",
       x = "Limb Volume (cm³)",
       y = "Frequency") +
  theme_light()
```
